---
title: "Final"
author: "Katie Miller"
format:
  html:
    toc: true
    toc-location: left
    code-fold: true
execute:
  message: false
  warning: false
---

# Problem 1

## Introduction

### Introduction Code

```{r}
# loading in packages
library(here)
library(tidyverse)
library(naniar)
library(GGally)
library(lmtest)
library(glmmTMB)
library(DHARMa)
library(rstatix)
library(flextable)
```

```{r}
# reading in data
rat <- read_csv(here("data","sev208_kratseedbank_20120213.txt"))
```

## Methods

### Methods Figures & Code
```{r}
gg_miss_var(rat) +
  theme_bw() 
```
**Figure 1. Visualizing missing data.** There are no missing values in our kangaroo rat dataset.

```{r}
ggplot(data = rat, aes(x = loc, y = seeds)) +
  geom_point() +
  theme_bw() +
  labs(x = "Mound location", y = "Total number of seeds")
```
**Figure 2. Exploratory data visualization.** There seems to be a correlation between mound location and number of seeds, further analysis should be conducted.

```{r}
# Kruskal-Wallis rank sum test
kruskal.test(seeds ~ loc, data = rat) 
```
**Table 1. Kruskal-Wallis rank sum test** At least two seed counts differ between the mound locations.

```{r}
# quantify the effect 
rat %>% 
  kruskal_effsize(seeds ~ loc)
```
**Table 2. Effect size for Kruskal Wallis.** The magnitude of the effect is small.

```{r}
ggplot(data = rat_summary, aes(x = loc, y = mean_l, color = loc)) +
  geom_point(size = 3) +
  geom_linerange(aes(ymin = mean_l - margin_l, ymax = mean_l + margin_l), linewidth = 1) +
  geom_jitter(data = rat, aes(x = loc, y = seeds), alpha = 0.3) +
 # scale_color_manual(values = c("Reference" = "#E57B33", "Calcium-treated" = "#039199")) +
  labs(x = "Mound location", y = "Total number of seeds") +
  theme_bw() +
  theme(legend.position = "none",
        text = element_text(family = "Times New Roman"),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12))
```
**Figure 3. Total number of seeds in each mound location.** The darker points represent the means with standard error bars and the lighter points represent the entire data set, The red represents ...

## Results

### Results Figures & Code

## Bibliography

# Problem 2

## Introduction

### Introduction Code

## Methods

### Methods Figures & Code

## Results

### Results Figures & Code

## Bibliography
