---
title: "Final"
author: "Katie Miller"
format:
  html:
    toc: true
    toc-location: left
    code-fold: true
execute:
  message: false
  warning: false
---

# Problem 1

## Introduction

The kangaroo rat (*Dipodomys sp.*) represents a genus of burrowing rodents which feed on various plant seeds. Kangaroo rats typically have a small circular home range of about 0.05 hectares where they collect large quantities of seeds for storage in their burrows (Schroder, 1979). Their mounds constitute small-scale disturbances which have been shown to alter the plant community by enhancing the surrounding seed bank (Alkon et al., 2010). Studies have shown that the kangaroo rat seed caching of several important range forage and browse species are essential to their regeneration (Sjoberg et al., 1984). Due to its relatively large impact on the ecosystem, the kangaroo rat is considered a keystone species (Chew, 2003). In this analysis we will be utilizing a dataset collected by Koontz and Simpson (2013) in order to understand how the quantity of seeds varies spatially between kangaroo rat mound locations; at the base of the mound (B), surrounding area of the mound (S), edge of Bouteloua eripoda (black grama) grass clumps (E), and the interspace between black grama grass clumps (I). If our hypothesis is supported, there will be a significant difference in the mean number of seeds between at least two locations.

### Introduction Code

```{r}
# loading in packages
library(here)
library(MASS)
library(tidyverse)
library(naniar)
library(GGally)
library(lmtest)
library(glmmTMB)
library(DHARMa)
library(rstatix)
library(flextable)
library(lme4)
library(MuMIn)
library(ggeffects)
```

```{r}
# reading in data
rat <- read_csv(here("data","sev208_kratseedbank_20120213.txt"))
```

## Methods

Koontz and Simpson (2013) collected data on kangaroo rat mounds in Sevilleta National Wildlife Refuge Long-term Ecological Research (LTER) site in Socorro County in central New Mexico to examine the effects of kangaroo rat mounds on surrounding plant communities. The resulting data was wrangled and uploaded to RStudio. We initially filtered the dataset to extract the variables most relevant to answering our research question: how does total seed number differ between kangaroo rat mound locations? The dataset is already cleaned with no missing data (Figure 1). Visually, there seems to be a relationship between the number of seeds versus mound location (Figure 2). Statistically, we used a Kruskal-Wallis rank sum test to determine that there is a significant difference in seed counts between at least two of the mound locations (Table 1, $x^2$(3) = 21.932, p-value = >0.001, $\alpha$ = 0.005). We calculated the effect size to quantify the effect, which was small (Table 2). This finding supports our hypothesis that there will be a significant difference in the mean number of seeds between at least two locations. To measure the variation in seed counts between different mound locations, we performed a Pairwise Wilcoxon Rank Sum Test. The greatest difference in seed counts occurred between the base of the mound and the inter-space between black grama grass and the mound (Table 3). 

### Methods Figures & Code
```{r}
gg_miss_var(rat) +
  theme_bw() 
```
**Figure 1. Visualizing missing data.** There are no missing values in our kangaroo rat dataset.

```{r}
ggplot(data = rat, aes(x = loc, y = seeds)) +
  geom_point() +
  theme_bw() +
  labs(x = "Mound location", y = "Total number of seeds")
```
**Figure 2. Exploratory data visualization.** There seems to be a correlation between mound location and number of seeds, further analysis should be conducted.

```{r}
# Kruskal-Wallis rank sum test
kruskal.test(seeds ~ loc, data = rat) 
```
**Table 1. Kruskal-Wallis rank sum test.** At least two seed counts differ between the mound locations.

```{r}
# quantify the effect 
rat %>% 
  kruskal_effsize(seeds ~ loc)
```
**Table 2. Effect size for Kruskal Wallis.** The magnitude of the effect is small.

```{r}
# Wilcoxon signed rank test
pairwise.wilcox.test(rat$seeds, rat$loc, p.adjust.method = "BH")
```
**Table 3. Pairwise Wilcoxon Rank Sum Test.** The biggest difference in seed counts is between the B (base of mound) and I (interspace between black grama grass and the mound).

## Results

We found that the number of seeds does vary by kangaroo rat mound location. Figure 3 compares the mean number of seeds in each mound location. There is a significant difference between mean seed counts at all locations. The highest mean number of seeds can be found at the base of the mound, and the lowest mean number of seeds can be found in the inter-space between black grama grass and the mound.

### Results Figures & Code

```{r}
# creating summary table
rat_summary <- rat %>% 
  group_by(loc) %>% 
  summarize(mean_seed = mean(seeds),
            sd_seed = sd(seeds),
            var_seed = var(seeds),
            count_seed = length(seeds),
            se_seed = sd_seed/sqrt(count_seed),
            margin_seed = qt(0.95, df = count_seed - 1) * se_seed)

# plotting relationship
ggplot(data = rat_summary, aes (y = mean_seed, x = loc, color = loc)) +
  geom_point() +
  geom_linerange(aes(ymin = mean_seed - margin_seed, ymax = mean_seed + margin_seed), linewidth = 0.5) +
  labs(x = "Mound location",
       y = "Mean number of seeds") +
  theme_bw() +
  theme(legend.position = "none",
        text = element_text(family = "Times New Roman"),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12)) 
```
**Figure 3. Mean number of seeds in each mound location.** The points represent the mean number of seeds with bars for standard error. The colors represent different mound locations, red is base of mound, green is surrounding, blue is edge of grass, and purple is the inter-space. 

## Bibliography

Alkon, P.U., et al. “The Composition of Seed Banks on Kangaroo Rat (Dipodomys Spectabilis) Mounds in a Chihuahuan Desert Grassland.” Journal of Arid Environments, 18 Apr. 2010, www.sciencedirect.com/science/article/pii/S0140196310000881?via%3Dihub#fig1. 

Chew, R.M., et al. “Keystone Effects of the Endangered Stephens’ Kangaroo Rat (Dipodomys Stephensi).” Biological Conservation, 1 July 2003, www.sciencedirect.com/science/article/pii/S0006320703001848?casa_token=uIBPgLsVrKMAAAAA%3A3xlSN3_dBYFf9nESVSU6R83eAquf6AhCdnoowrJ_qsC7eSRScAeAmCcWXRmTP65lBxKkSvCCyRI. 

Koontz, T. and H. Simpson. 2013. Effects of Kangaroo Rat Mounds on Seed Banks of Grass and Shrublands at the Sevilleta National Wildlife Refuge, New Mexico (2001) ver 102459. Environmental Data Initiative. https://doi.org/10.6073/pasta/115b8095a5e82028869a8c56eff14597 

Schroder, Gene. “Foraging Behavior and Home Range Utilization of the Bannertail Kangaroo Rat.” Ecology, 1 Aug. 1979, esajournals.onlinelibrary.wiley.com/doi/abs/10.2307/1936601. 

Sjoberg, Diana. “Kangaroo Rats.” University of Arizona, Feb. 1984, journals.uair.arizona.edu/index.php/rangelands/article/viewFile/11849/11122. 


# Problem 2

## Introduction

Vegetation composition shifts are occurring due to climate change and tundras are particularly at risk from the expansion of woody shrubland (Mekonnen, 2021).
Plant phenology depends on environmental markers to occur, and changes in climate can alter the lifecycle and reproduction of plants, thus altering plant communities and species composition as a whole (Cleland, 2007). The expansion of shrub combined with the lower productivity of tundra plants due to climate change increases the risk of local extinction (Heidi and Luoto, 2016). 
In this analysis, we will be utilizing a data set collected by Seaver (2022) to examine how seed count varies with plant species, total number of inflorescences, and open versus shrub plot type. We are also interested in understanding whether there is a simpler model explaining seed count.


### Introduction Code

```{r}
# reading in data
seed <- read_csv(here("data", "knb-lter-nwt.297.1", "shrubstudy_seed_ctwt.ms.data.csv"))
```

## Methods

Seaver (2022) collected data on flowering phenology, seed counts, and pollinator visitation rates in shrub and open plots in Niwot Ridge between 2019 - 2021. We initially filtered this dataset to extract the variables most relevant to answering our research question: how does seed count vary with plot type, plant species, and total number of inflorescences? If our hypothesis is supported, these variables will predict seed count. We are also interested in examining whether there is a simpler model to explain variation in seed count. After initial wrangling and filtering, we began our analysis by creating a visualization of the missing data in our entire data subset (Figure 1). There is some missingness in the total number of seeds, the weight of seeds, and the distance from shrub. 

To determine how seed count varies with plot type, plant species, and total number of inflorescences, we created a null model and multiple ful models using different formulas of linear regression. We ranked our models by AIC and the best fitting model was the negative binomial generalized linear model (Table 1).  An assumption check for the selected model shows normally distributed residuals, however, some quantile deviations were detected. We are choosing to move forward with it as it the simplest of all model options. The AIC of the model is lower than that of the null model, therefore we can say that plot type, plant species, and total number of inflorescences do predict seed count.

There may be a simpler model which explains variance in seed counts. We created a pair plot of all of our variables of interest compared against each other to better understand relationships between our variables (Figure 3). There seems to be a strong correlation between number of seeds and weight of seeds, total number of inflorescences, and plot aspect. We then created more full models with these three variables as predictors and used AIC to compare them to the previously selected model and the null model. The best fitting model we found to explain the total number of seeds was the weight of seeds and total number of inflorescences (Table 2).  Our simpler model shows normally distributed residuals with quantile deviations in the residuals vs. predicted plot.

### Methods Figures & Code

```{r}
# visualizing missing data
gg_miss_var(seed)
```
**Figure 1. Visualizing missing data.** Our response variable, total number of seeds, has around 100 missing values.

```{r}
# creating a subset with variables of interest
seed_subset <- seed %>% 
  select(aspect, treatment, species, total_nr_infl, nr_seeds, 
         dist_fr_shr, wt_seeds_mg)
```

```{r}
# model selection: creating null model and testing different fits

null_seedmod <- glm.nb(nr_seeds ~ 1, data = seed_subset)

seedmod1 <- glm(nr_seeds ~ species + treatment + total_nr_infl, data = seed_subset, family = "poisson")

seedmod2 <- glm.nb(nr_seeds ~ species + treatment + total_nr_infl, data = seed_subset)

seedmod3 <- glmer(nr_seeds ~ species + treatment + total_nr_infl + (1|aspect), data = seed_subset, family = "poisson")

seedmod4 <- glmer.nb(nr_seeds ~ species + treatment + total_nr_infl + (1|aspect), data = seed_subset)

MuMIn::model.sel(null_seedmod, seedmod1, seedmod2, seedmod3, seedmod4)
```
**Table 1. Model selection.** Of our models, the Negative Binomial Generalized Linear Model fit the best. 

```{r}
# plotting simulated residuals for the the best fitting model
plot(simulateResiduals(seedmod2))
```
**Figure 2. Assumptions check for selected model.** The black points represent residuals. The red lines depict patterns in the residuals. 

```{r}
# simpler model?
seed_subset %>% 
  ggpairs(cardinality_threshold = 19)
```
**Figure 3. Pairs plot.** Allows us to see if there is a correlation between variables. 

```{r}
# simpler model to explain seed count?
seedmod5 <- glm.nb(nr_seeds ~ wt_seeds_mg, data = seed_subset)

seedmod6 <- glm.nb(nr_seeds ~ wt_seeds_mg + total_nr_infl, data = seed_subset)

seedmod7 <- glm.nb(nr_seeds ~ wt_seeds_mg + total_nr_infl + aspect, data = seed_subset)

MuMIn::model.sel(seedmod2, seedmod5, seedmod6, seedmod7)
```
**Table 2. Simpler model selection.** Model 6 containing seed weight and total number of inflorescences is the simplest.

```{r}
plot(simulateResiduals(seedmod6))
```
**Figure 4. Check of assumptions for simplified model.** The black points represent residuals. The red lines depict patterns in the residuals. 


## Results

Our hypothesis that plot type, plant species, and total number of inflorescences predict seed count is supported due to the fact that the AIC of the model is lower than that of the null model. Our models of predictions diverge at high number of inflorescence, this is likely due to large amount of missingness in the dataset. At low values, the model seems to fit well to the data with smaller error bars (Figure 5).

This model is suitable, however, there is a simpler model to explain variation in the total number of seeds. The best fitting model we found was the weight of seeds and total number of inflorescences. Comparatively, the original model had fewer correlated variables and a larger standard error (Table 3, Table 4). This model also makes biological sense; the total number of inflorescences would affect seed quantity because the number of seeds produced by a plant is dependent upon the number of flowers produced and pollinated. Plants which produce heavier seeds are likely bigger and healthier, therefore likely to produce a greater quantity of seeds as well. 

fewer correlated variables and a larger standard error ()


### Results Figures & Code

```{r}
#
seedmod2 %>% 
  as_flextable()
```
**Table 3. Summary of original model.**

```{r}
# visualizing predictions for model 2

# removing NAs
seed.na <- seed %>% 
  select(treatment, species, total_nr_infl, nr_seeds) 

seed.filtered <- na.omit(seed.na)

# creating model predictions 
predictions <- ggpredict(seedmod2, terms = c("total_nr_infl", "species", "treatment")) %>% 
  rename(treatment = facet,
         species = group)

# graphing predictions
ggplot(seed.filtered, aes(x = total_nr_infl, y = nr_seeds, color = treatment, na.rm = TRUE)) +
  geom_point(alpha = 0.4) +
  geom_line(data = predictions, aes(x = x, y = predicted, color = treatment)) +
  geom_ribbon(data = predictions, aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high, fill = treatment), alpha = 0.2) +
  scale_fill_manual(values = c("shrub" = "red", "control" = "blue")) +
  scale_color_manual(values = c("shrub" = "red", "control" = "blue")) +
  facet_wrap(~ species) +
  labs(x = "Number of inflorescences", y = "Number of seeds") +
  theme_bw() +
   theme(text = element_text(family = "Times New Roman"),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12)) 
```
**Figure 5. Visualizing model predictions.** Lines predicted number of seeds by inflorescences in the open plot (blue line) in the shub plot (red line). The shaded areas represent the confidence intervals. 

```{r}
#
seedmod6 %>% 
  as_flextable()
```
**Table 4. Summary of simplified model.**

## Bibliography

Cleland, E. et al. “Shifting Plant Phenology in Response to Global Change.” Trends in Ecology & Evolution, 2 May 2007, www.sciencedirect.com/science/article/pii/S0169534707001309. 

Heidi K, and Miska Luoto. “Arctic Shrubification Mediates the Impacts of Warming Climate on Changes to Tundra Vegetation.” Environmental Research Letters, 25 Nov. 2016, iopscience.iop.org/article/10.1088/1748-9326/11/12/124028/meta. 

Mekonnen, Zelalem et al. “IOPscience.” Environmental Research Letters, 23 Apr. 2021, iopscience.iop.org/article/10.1088/1748-9326/abf28b/meta. 

Seaver, M. 2022. Individual and community flowering phenology, seed counts and pollinator visitation rates in shrub and open plots across Niwot Ridge, 2019 - 2021. ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/edc4ab2faf7dd96cd1deac1544d5f2b9 

