---
title: "Final"
author: "Katie Miller"
format:
  html:
    toc: true
    toc-location: left
    code-fold: true
execute:
  message: false
  warning: false
---

# Problem 1

## Introduction

### Introduction Code

```{r}
# loading in packages
library(here)
library(tidyverse)
library(naniar)
library(GGally)
library(lmtest)
library(glmmTMB)
library(DHARMa)
library(rstatix)
```

```{r}
# reading in data
rat <- read_csv(here("data","sev208_kratseedbank_20120213.txt"))
```

## Methods

### Methods Figures & Code
```{r}
gg_miss_var(rat) +
  theme_bw() 
```
**Figure 1. Visualizing missing data.** There are no missing values in our kangaroo rat dataset.

```{r}
ggplot(data = rat, aes(x = loc, y = seeds)) +
  geom_point() +
  theme_bw() +
  labs(x = "Mound location", y = "Total number of seeds")
```
**Figure 2. Exploratory data visualization.** There seems to be a correlation between mound location and number of seeds, further analysis should be conducted.

```{r}
# Kruskal-Wallis rank sum test
kruskal.test(seeds ~ loc, data = rat)
```
**Table 1. Kruskal-Wallis rank sum test** At least two seed counts differ between the mound locations.

```{r}
# quantify the effect 
rat %>% 
  kruskal_effsize(seeds ~ loc)
```
**Table 2. Effect size for Kruskal Wallis.** The magnitude of the effect is small.

## Results

### Results Figures & Code

## Bibliography

# Problem 2

## Introduction

### Introduction Code

## Methods

### Methods Figures & Code

## Results

### Results Figures & Code

## Bibliography
